---
title: Homework 5
author: Daniel Dulaney
date: November 7, 2020
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(here)
library(ggrepel)
library(ISLR)
library(e1071)
```

#(9.3)

##(a)

```{r}
df_9.3 <- tribble(
  ~obs, ~x1, ~x2, ~y,
  1, 3, 4, "Red",
  2, 2, 2, "Red",
  3, 4, 4, "Red",
  4, 1, 4, "Red",
  5, 2, 1, "Blue",
  6, 4, 3, "Blue",
  7, 4, 1, "Blue"
)

points_9.3 <- df_9.3 %>% 
  ggplot(aes(x1, x2, color = y)) +
  geom_point(size = 6) +
  geom_label_repel(aes(label = obs)) +
  scale_color_manual(values = c("Blue", "Red"))

points_9.3
```

##(b)

Based on the plot, we know we can construct a line that will separate the classes perfectly.

The maximal margin classifier will be a line between Blue points 5 (2, 1) and 6 (4, 3), and red points 2 (2, 2) and 3 (4, 4). The line will need to pass through the two points (4, 3.5) and (2, 1.5).

The slope of the line equals $\frac{3.5-1.5}{4-2} = 1$ and the intercept equals 1.5-2 = -0.5

```{r}
points_9.3 +
  geom_abline(intercept = -.5,
              slope = 1,
              size = 3)
```

##(c)

0.5 - $X_1$ + $X_2$ > 0

##(d)

The line passing through Red points 2 and 3 has slope 1 and intercept -1, and the line passing through Blue points 5 and 6 has slope 1 and intercept 0

```{r}
points_9.3 +
  geom_abline(intercept = -.5,
              slope = 1,
              size = 3) +
  geom_abline(intercept = -1,
              slope = 1,
              size = 2,
              linetype = "dashed") +
  geom_abline(intercept = 0,
              slope = 1,
              size = 2,
              linetype = "dashed")
```

##(e)

```{r}
points_9.3 +
  geom_abline(intercept = -.5,
              slope = 1,
              size = 3) +
  geom_segment(aes(x = 2,
                   y = 2,
                   xend = 2,
                   yend = 1),
               size = 3) +
  geom_segment(aes(x = 4,
                   y = 3,
                   xend = 4,
                   yend = 4),
               size = 3)
```

##(f)

Since observation 7 is outside of the margin, which is the area between the dotted lines, moving it wouldn't affect the maximal margin hyperplane.

##(g)

```{r}
points_9.3 +
  geom_abline(intercept = -1.1,
              slope = 1,
              size = 3)
```

$-1.1 - X_1 + X_2 > 0$

##(h)

```{r}
df_9.3 %>% 
  rbind(c(8, 5, 1, "Red")) %>% 
  ggplot(aes(x1, x2, color = y)) +
  geom_point(size = 6) +
  geom_abline(intercept = -1.1,
              slope = 1,
              size = 3) +
  geom_label_repel(aes(label = obs)) +
  scale_color_manual(values = c("Blue", "Red"))
```

#(9.7)

```{r}
auto <- as_tibble(Auto)
```

##(a)

```{r}
mpg_median <- median(auto$mpg)

auto <- auto %>% 
  mutate(high_mileage = ifelse(mpg > mpg_median, 1, 0))
```

##(b)

```{r}
tune_linear <- tune(svm, high_mileage ~ ., data = auto, kernel = "linear", ranges = list(cost = c(.01, .1, 1, 5, 10, 100)))

summary(tune_linear)
```

Errors for various costs in the table above. The lowest error was at cost = 1.

##(c)

```{r}
tune_poly <- tune(svm, high_mileage ~ ., data = auto, kernel = "polynomial", ranges = list(cost = c(0.1, .5, 1, 3, 15), degree = c(2, 3, 4, 5)))

summary(tune_poly)
```

Best cost for type polynomial is at cost = 15.

```{r}
tune_radial <- tune(svm, high_mileage ~ ., data = auto, kernel = "radial", ranges = list(cost = c(0.1, .5, 1, 3, 15), degree = c(2, 3, 4, 5)))

summary(tune_radial)
```

Best cost for type radial is at cost = 15.

##(d)

```{r}
svm_linear <- svm(high_mileage ~ ., data = auto, kernel = "linear", cost = 1)

plot(svm_linear, auto, as.formula(mpg~cylinders))
```

This code follows what is in the book but it's not producing the plot here. Not sure what the issue is.

#(10.2)

Will use the `hclust()` function from an earlier chapter to create these trees.

##(a)

```{r}
matrix_10.2 <- as.matrix(tribble(
  ~a, ~b, ~c, ~d,
  0, 0.3, 0.4, 0.7,
  0.3, 0, 0.5, 0.8,
  0.4, 0.5, 0, 0.45,
  0.7, 0.8, 0.45, 0
))

dist_10.2 <- as.dist(matrix_10.2)

plot(hclust(dist_10.2, method = "complete"))
```

##(b)

```{r}
plot(hclust(dist_10.2, method = "single"))
```

##(c)

They are split into (a, b) and (c, d)

##(d)

They are split into (d) and (c, b, a)

##(e)

```{r}
plot(hclust(dist_10.2, method = "complete"), 
     labels = c("b", "a", "d", "c"))
```

#(10.3)

##(a)

```{r}
df_10.3 <- tribble(
  ~obs, ~x1, ~x2,
  1, 1, 4,
  2, 1, 3,
  3, 0, 4,
  4, 5, 1,
  5, 6, 2,
  6, 4, 0
)

df_10.3 %>% 
  ggplot(aes(x1, x2)) +
  geom_point()
```

##(b)

```{r}
df_10.3 <- df_10.3 %>% 
  mutate(cluster = sample(2, 6, replace = TRUE))

df_10.3
```

Clusters in last column.

##(c)

```{r}
# calculate centroids
clust1_x1 <- df_10.3 %>% 
  filter(cluster == 1) %>% 
  summarise(mean(x1)) %>% 
  pull()

clust1_x2 <- df_10.3 %>% 
  filter(cluster == 1) %>% 
  summarise(mean(x2)) %>% 
  pull()

clust2_x1 <- df_10.3 %>% 
  filter(cluster == 2) %>% 
  summarise(mean(x1)) %>% 
  pull()

clust2_x2 <- df_10.3 %>% 
  filter(cluster == 2) %>% 
  summarise(mean(x2)) %>% 
  pull()
```

```{r}
df_10.3 %>% 
  ggplot() +
  geom_point(aes(x1, x2, color = as.factor(cluster)),
             data = df_10.3,
             size = 5) +
  geom_point(aes(x1, x2),
             data = tribble(~x1, ~x2,
                            clust1_x1, clust1_x2),
             color = "red",
             size = 5,
             shape = "triangle") +
  geom_point(aes(x1, x2),
             data = tribble(~x1, ~x2,
                            clust2_x1, clust2_x2),
             color = "blue",
             size = 5,
             shape = "triangle") +
  scale_color_manual(values = c("red", "blue"))
```

##(d)

```{r}
df_10.3 <- df_10.3 %>% 
  mutate(center_clust1_x1 = clust1_x1,
         center_clust1_x2 = clust1_x2,
         center_clust2_x1 = clust2_x1,
         center_clust2_x2 = clust2_x2,
         dist_to_centerclust1 = sqrt((center_clust1_x1 - x1)^2 + (center_clust1_x2 - x2)^2),
         dist_to_centerclust2 = sqrt((center_clust2_x1 - x1)^2 + (center_clust2_x2 - x2)^2),
         new_cluster = ifelse(dist_to_centerclust1 < dist_to_centerclust2, 1, 2))

df_10.3 
```

```{r}
df_10.3 %>% 
  ggplot() +
  geom_point(aes(x1, x2, color = as.factor(new_cluster)),
             data = df_10.3,
             size = 5) +
  geom_point(aes(x1, x2),
             data = tribble(~x1, ~x2,
                            clust1_x1, clust1_x2),
             color = "red",
             size = 5,
             shape = "triangle") +
  geom_point(aes(x1, x2),
             data = tribble(~x1, ~x2,
                            clust2_x1, clust2_x2),
             color = "blue",
             size = 5,
             shape = "triangle") +
  scale_color_manual(values = c("red", "blue"))
```

##(e)

```{r}
# calculate centroids
clust1new_x1 <- df_10.3 %>% 
  filter(new_cluster == 1) %>% 
  summarise(mean(x1)) %>% 
  pull()

clust1new_x2 <- df_10.3 %>% 
  filter(new_cluster == 1) %>% 
  summarise(mean(x2)) %>% 
  pull()

clust2new_x1 <- df_10.3 %>% 
  filter(new_cluster == 2) %>% 
  summarise(mean(x1)) %>% 
  pull()

clust2new_x2 <- df_10.3 %>% 
  filter(new_cluster == 1) %>% 
  summarise(mean(x2)) %>% 
  pull()
```

##(f)

```{r}

```

#(10.9)

##(a)

```{r}
usa_arrests <- as_tibble(USArrests %>% 
                           rownames_to_column(var = "state"))
```

```{r}
hc_complete <- hclust(dist(usa_arrests), method = "complete")

plot(hc_complete)
```

##(b)

```{r}
usa_arrests <- usa_arrests %>% 
  mutate(cluster = cutree(hc_complete, 3))

# show each cluster individually
usa_arrests %>% 
  filter(cluster == 1)

usa_arrests %>% 
  filter(cluster == 2)

usa_arrests %>% 
  filter(cluster == 3)
```

##(c)

```{r}
usa_arrests_scaled <- scale(usa_arrests %>% 
                              column_to_rownames("state") %>% 
                              select(-cluster))

hc_scaled <- hclust(dist(usa_arrests_scaled), method = "complete")

plot(hc_scaled)
```

##(d)

#(10.10)

##(a)

```{r}
mat_10.10 <- matrix(nrow = 60, ncol = 50)

for (i in 1:20) {
  mat_10.10[i, ] <- c(rnorm(n = 50, mean = 0))
}

for (i in 21:40) {
  mat_10.10[i, ] <- c(rnorm(n = 50, mean = 10))
}

for (i in 41:60) {
  mat_10.10[i, ] <- c(rnorm(n = 50, mean = 20))
}

# df_10.10 <- as_tibble(mat_10.10) %>% 
#   mutate(cluster = c(rep(1, 20), rep(2, 20), rep(3, 20)))
# 
# df_10.10
```

##(b)

```{r}
pca_out <- prcomp(mat_10.10)

summary(pca_out)
```

##(c)

```{r}
km_out <- kmeans(mat_10.10, 3, nstart = 20)

km_out$cluster
```

As in the dataset, points are clustered into different classes 20 by 20 by 20. They are perfectly learned here as I gave them a large mean separation.

##(d)

```{r}
km_out2 <- kmeans(mat_10.10, 2, nstart = 20)

km_out2$cluster
```

The second class which was previously "2" is now a part of class "1", while the other class (previously 3) is now all 2.

##(e)

```{r}
km_out4 <- kmeans(mat_10.10, 4, nstart = 20)

km_out4$cluster
```

The cluster which was originally 2, rows 21:40 from the data, is split into two different clusters now. The other two remain the same as original, perfectly separated into their own clusters.

##(f)

```{r}
km_outpca <- kmeans(pca_out$x[, 1:2], 3, nstart = 20)

km_outpca$cluster
```

Like we saw at the start, split perfectly into 3 clusters.

##(g)

```{r}
mat_10.10_scaled <- scale(mat_10.10)

km_outscaled <- kmeans(mat_10.10_scaled, 3, nstart = 20)

km_outscaled$cluster
```

They are still perfectly separated.





